---
title: "Hypothesis Testing"
author: "Yang Xi"
date: "09 Jul, 2016"
output:
  html_document:
    toc: 1
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

<br>

### Shapiro-Wilk Test: Exam Normal Distribution
**Null hypothesis**: normal
```{r}
set.seed(1)
shapiro.test(rnorm(100, mean = 5, sd = 3)) # p-value=0.99 > 0.05 => normally distributed
shapiro.test(runif(100, min = 2, max = 4)) # very small p-value => NOT normally distributed
```

<br>

### F-test: Whether the Variance(s) of Two Samples are Equal
**Assumption**: sample from normal distributed population <br>
**Null hypothesis**: the two variances are equal
```{r}
set.seed(1)
n <- 10000
V1 <- runif(n, min=0, max=1)
V2 <- runif(n, min=0, max=1)
var.test(V1, V2, conf.level=0.95, alternative="two.sided") # p-value=0.4184>0.05 => no significant difference

set.seed(1)
V1 <- runif(n, min=0, max=1)
V2 <- runif(n, min=0, max=10)
var.test(V1, V2, conf.level=0.95, alternative="two.sided") # p-value=2.2e-16 => significant difference
var.test(V1, V2, conf.level=0.95, alternative="less")  # p-value=2.2e-16 => V1 significant less then V2
var.test(V1, V2, conf.level=0.95, alternative="greater")  # p-value=1

```

### t-test: Whether the Mean(s) of Two Groups are Equal
**Assumption**: sampled from normal distributions with equal variances<br>
**Null hypothesis**: tow means are equal.
```{r}
x <- c(1,2,3,4,5,6,7,8,9,10)
y <- c(2,3,4,5,6,7,8,9)

t.test(x, y, alternative="two.sided")
```

<br>

### Chi-Square Test of Independence between Two Categorical Variables
**Null hypothesis**: the two variables are independent
```{r}
vSize <- ifelse(iris$Sepal.Length < median(iris$Sepal.Length), "small", "big")
vSpecies <- iris$Species

chisq.test(vSize, vSpecies) # small p-value => significant relationship (not independent)
```

<br>
