---
title: "Probability Distribution"
author: "Yang Xi"
date: "09 Jul, 2016"
output:
  html_document:
    toc: 1
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

<br>

# Discrete Distribution
```{r}
x <- c(0, 1, 2)
Px <- c(0.25, 0.5, 0.25)

Ex <- weighted.mean(x, Px)
Var <- sum((x-Ex)^2*Px)
Std <- sqrt(Var)

print(paste0('mean=',Ex, ' variance=', Var, ' standard deviation=',sprintf("%0.5f", Std)))
```

<br>

### Binomial Distribution
Probability of `x` successes in `n` trials, given the probability of success `P` on each trail (success or fail).

* Label manufactured items as acceptable or defective
* Survey responses of yes or no
```{r}
x <- 4
n <- 12
P <- 0.2

Ex <- n*P
Var <- n*P*(1-P)
Std <- sqrt(Var)
print(paste0('mean=',Ex, ' variance=', Var, ' standard deviation=',sprintf("%0.5f", Std)))

Px <- dbinom(x, size=n, prob=P)
Fx <- pbinom(x, size=n, prob=P, lower.tail = TRUE)

print(paste0('PDF=',Px,' CDF=',Fx))
```

<br>

### Hypergeometric Distribution
The finite population contains `m` white balls and `n` black balls. If draw `k` balls without replacement, what is the probability of draw `x` white balls.
```{r}
Px <- dhyper(x=2, m=4, n=6, k=3)
Fx <- phyper(x, m=4, n=6, k=3, lower.tail = TRUE)
print(paste0('PDF=',Px,' CDF=',Fx))
```

<br>

### Poisson Distribution
Probability of `x` occurrence per unit(time), given the expected number of occurrence per unit(time) `lambda`.
```{r}
x <- 2
lambda <- 0.5

Ex <- lambda
Var <- lambda
Std <- sqrt(lambda)
print(paste0('mean=',Ex, ' variance=', Var, ' standard deviation=',sprintf("%0.5f", Std)))

Px <- dpois(x, lambda)
Fx <- ppois(x, lambda, lower.tail = TRUE)
print(paste0('PDF=',sprintf("%0.5f", Px),' CDF=',sprintf("%0.5f", Fx)))
```

<br>

# Continuous Distribution
### Uniform Distribution
```{r}
x <- 0.4
a <- 0
b <- 1

Ex <- (a+b)/2
Var <- (b-a)^2/12
Std <- sqrt(Var)
print(paste0('mean=',Ex, ' variance=', sprintf("%0.5f", Var), ' standard deviation=',sprintf("%0.5f", Std)))


Px <- dunif(x, a, b)
Fx <- punif(x, a, b, lower.tail = TRUE)
print(paste0('PDF=',sprintf("%0.5f", Px),' CDF=',sprintf("%0.5f", Fx)))
```

<br>

### Normal Distribution
```{r}
x <- 8.6
m <- 8
s <- 5

Px <- dnorm(x, m, s)
Fx <- pnorm(x, m, s, lower.tail = TRUE)
print(paste0('PDF=',sprintf("%0.5f", Px),' CDF=',sprintf("%0.5f", Fx)))
```

<br>

### Exponential Distribution
Probability of time between two occurrence is `x`, given the number of occurrence per unit time `lambda`.<br>
Commonly used to model the length of time between two occurrences,
```{r}
x <- 0.05
lambda <- 15

Ex <- 1/lambda
Var <- 1/lambda^2
Std <- sqrt(Var)
print(paste0('mean=',sprintf("%0.5f", Ex), ' variance=', sprintf("%0.5f", Var), ' standard deviation=',sprintf("%0.5f", Std)))

Px <- dexp(x, lambda)
Fx <- pexp(x, lambda, lower.tail = TRUE)
print(paste0('PDF=',sprintf("%0.5f", Px),' CDF=',sprintf("%0.5f", Fx)))

```

<br>

# Covariance, Correlation and Independence
The covariance / correlation measures the strength of the linear relationship between two variable.

* If two random variable are statistically independent, the covariance between them is 0.
* **NOTE**: the converse is not necessarily true.
```{r}
x <- c(1,2,3,4,5)
y <- c(2,1,5,3,2)
cov(x, y, method="pearson")
cor(x, y, method="pearson")
```

<br>



