---
title: "Probability Distribution"
author: "Yang Xi"
date: "09 Jul, 2016"
output:
  html_document:
    toc: 1
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

<br>

# Sampling Distribution of **Sample Mean**
If population is **normal** with mean `u` and standard deviation `sigma`, the sampling distribution of sample mean is also normally distributed.<br>
`n` is the sample size.
```{r}
u <- 10
sigma <- 2
n <- 100

mean_sample_mean <- u
std_sample_mean <- sigma/sqrt(n)

print(paste0('mean of sample mean=',mean_sample_mean, '  std of sample mean=', std_sample_mean))
```

<br>

### Finite Population Correction
Apply the finite population correction if:

* sampling is **without** replacement
* sample is is large relative to population ($n>0.05*N$)
```{r}
u <- 10
sigma <- 2
n <- 100
N <- 1000

std_sample_mean = sigma/sqrt(n)*sqrt((N-n)/(N-1))
print(paste0('corrected std of sample mean=', sprintf("%0.5f", std_sample_mean)))
```

<br>

### Central Limited Theorem
Even if the population is **not normal**, sample means from the population will be approximately normal as long as the sample size is large enough (n > 25).

<br>

# Sampling Distribution of **Sample Proportion**
Given the population is a basket of white/black balls, with probability of `P` being white.<br>
The **sample proportion** has binomial distribution, but can be approximated by a normal distribution when $nP(1-P)>9$.
```{r}
P <- 0.4
n <- 200

mean_sample_proportion <- P
std_sample_proportion <- sqrt(P*(1-P)/n)

print(paste0('mean of sample proportion=',mean_sample_proportion, '  std of sample proportion=', sprintf("%0.5f", std_sample_proportion)))
```

<br>

# Sampling Distribution of **Sample Variance**
Draw samples with size `n` from population with standard deviation `sigma`:
```{r}
sigma <- 2
n <- 100

mean_sample_variance <- sigma^2

# If population is normal
var_sample_variance <- 2*sigma^4/(n-1)

print(paste0('mean of sample variance=',mean_sample_variance, '  var of sample variance=', sprintf("%0.5f", var_sample_variance)))

```

<br>

## Chi-Square Distribution
Let `s` be the sample standard deviation.<br>
If population distribution is normal, then $(n-1)s^2/sigma^2$ has a chi-square distribution with n-1 degree of freedom.

<br>

### Example
An air-con product must hold a specific temperature with little variation.<br>
Requirement calls for a standard deviation of no more than 4 degrees (or variance of 16 degrees2).<br>
A sample of 14 air-cons is to be tested.<br>
What is the **upper limit of the sample variance**, such that the probability of exceeding this limit is less than 0.05?
```{r}
sigma <- 4
n <- 14
alpha <- 0.05

X2 <- qchisq(alpha, df=n-1, lower.tail=FALSE)

s2 <- X2*sigma^2/(n-1)
print(paste('the upper limit of the sample variance is', sprintf("%0.2f", s2)))

```

<br>
