---
title: "Dynamic Harmonic Regression"
author: "Yang Xi"
date: "17 Jan, 2018"
output:
  html_document:
    toc: 1
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

<br>

# 8. Dynamic Harmonic Regression
### 8.1 Definition
Besides seasonal models, Fourier transformation has been a traditional method to handle periodicity in signal.

<br>
Let
$$ s_k(t)=sin(\frac{2\pi kt}{m}) $$
$$ c_k(t)=cos(\frac{2\pi kt}{m}) $$
$$ y_0=\beta_0+\sum_{k=1}^K[\alpha_k s_k (t)+\gamma_k c_k (t)]+e_t $$
where,<br>
$m=$ seasonal period<br>
$\alpha_k,\gamma_k=$ regression coefficients<br>
$e_t$ can be modeled as a non-seasonal ARIMA process

<br>
Note that, the Fourier model assumes that seasonal pattern **does not change over time**, while a seasonal ARIMA model allows the pattern to evolve over time.

<br>

### 8.2 Example: Monthly Sales Data of an Item
**8.2.1 Parameter Selection**<br>
Larger *K* value will result in more period patterns extracted. The following figures compare the forecasts with *K=1* and *K=6* on this monthly sales data respectively, with external regressors (*c* and *d*):
```{r, message=FALSE}
library(dplyr)
library(fpp2)

load('../../data/(2017) Time Series/time_series_monthly.rdata')

dfItem10 <- dfSalesMonthly %>%
  filter(item_id == 10 & first_date >='2008-01-01' & first_date <='2013-12-31') %>%
  as.data.frame() %>%
  mutate(coverage = (100-stock_out_pct)/100) %>%
  select(c(date=first_date, sales_qty, discount_pct, coverage))

# Train test data index
## 2008 - 2012 as training, 2013 as testing
tp10 <- 60

# Smooth outliers in training data
outlier_smooth <- function(df_train) {
  # Type 1: 0 sales, as log 0 cannot be handled.
  sales <- df_train$sales_qty
  outlier1 <- sales==0
  
  # Type 2: coverage is 0, as log 0 cannot be handled.
  outlier2 <- df_train$coverage==0

  # Type 3: low discount while high sales, cannot be explained by model
  discount <- df_train$discount_pct
  low_discount <- mean(discount)
  low_discount_i <- discount < low_discount
  high_sales <- mean(sales)+2*sd(sales)
  high_sales_i <- sales > high_sales
  outlier3 <- high_sales_i & low_discount_i
  
  # Adjust sales
  outliers <- outlier1 | outlier2 | outlier3
  
  df_train$sales_qty[outliers] <- NA 

  return(df_train)
}

dfItem10[1:tp10,] <- outlier_smooth(dfItem10[1:tp10,])

# Transform Variables
## Function to get a straight line connecting two points
get_line <- function(xa, ya, xb, yb, v) v*(ya-yb)/(xa-xb) + (yb*xa - ya*xb)/(xa-xb)

## Four-stage dm curve based on prior knowledge
discount_curve <- function(discount) {
  x1 <- 0
  y1 <- 1
  x2<- 5
  y2 <- 1.3
  x3 <- 30
  y3 <- 1.9
  x4 <- 40
  y4 <- 3
  x5 <- 100
  y5 <- 3
  
  ifelse(discount < x2, get_line(x1, y1, x2, y2, discount),
         ifelse(discount < x3, get_line(x2, y2, x3, y3, discount),
                ifelse(discount < x4, get_line(x3, y3, x4, y4, discount),
                       get_line(x4, y4, x5, y5, discount))))
}

dfItem10 <- dfItem10 %>%
  mutate(cm = coverage,
         dm = discount_curve(discount_pct),
         Y = log(sales_qty),
         Xc = log(cm),
         Xd = log(dm))

# Covert to ts
ts10 <- select(dfItem10, -date)
row.names(ts10) <- dfItem10$date
ts10 <- ts(ts10, start=c(2008, 1), frequency=12)
ts10s <- ts10[,'sales_qty']
#autoplot(ts10, facets=T)

# Separate train test sets
# 2008 - 2012 as training, 2013 as testing
ts10_train <- window(ts10, end = c(2013, 0))
ts10_test <- window(ts10, start = c(2013, 1))
ts10s_test <- window(ts10s, start = c(2013, 1))
ts10s_actual <- as.numeric(ts10s_test)
```

<br>
**_K=1_**
```{r}
# Function for inverse transformation
Y_inv <- function(fc) {
  fc$x <- exp(fc$x)
  fc$mean <- exp(fc$mean)
  fc$lower <- exp(fc$lower)
  fc$upper <- exp(fc$upper)
  return(fc)
}

# Forecast Accuracy
rmse_e <- function(e) {
  round(sqrt(mean(e^2, na.rm=T)), digits=2)
}


rmse <- function(pred, actual) {
  rmse_e(pred-actual)
}

mase <- function(pred, actual) {
  l <- length(pred)
  round(sum(abs(pred-actual))/sum(abs(diff(actual)))*(l-1)/l, digits=2)
}

# K = 1
fit_dhx1 <- auto.arima(ts10_train[,'Y'], xreg = cbind(ts10_train[,c('Xc', 'Xd')], fourier(ts10_train[,'Y'], K = 1)), seasonal=F)
fc_dhx1 <- forecast(fit_dhx1, xreg = cbind(ts10_test[,c('Xc', 'Xd')], fourier(ts10_train[,'Y'], K = 1, h=12)), level=c(0,0))
fci_dhx1 <- Y_inv(fc_dhx1)
autoplot(fci_dhx1) + autolayer(ts10s_test)
```

<br>
**_K=6_**
```{r}
fit_dhx6 <- auto.arima(ts10_train[,'Y'], xreg = cbind(ts10_train[,c('Xc', 'Xd')], fourier(ts10_train[,'Y'], K = 6)), seasonal=F)
fc_dhx6 <- forecast(fit_dhx6, xreg = cbind(ts10_test[,c('Xc', 'Xd')], fourier(ts10_train[,'Y'], K = 6, h=12)), level=c(0,0))
fci_dhx6 <- Y_inv(fc_dhx6)
autoplot(fci_dhx6) + autolayer(ts10s_test)
```

<br>
The optimal *K* value can be selected by minimizing AICc:
```{r}
# Optimal K
get_dhx_AICc <- function(K) {
  fit_dhx <- auto.arima(ts10_train[,'Y'], xreg = cbind(ts10_train[,c('Xc', 'Xd')], fourier(ts10_train[,'Y'], K = K)), seasonal=F)
  return(fit_dhx$aicc)
}

Ks <- data.frame(K=seq(1, 6))
Ks$AICc <- sapply(Ks$K, get_dhx_AICc)
Ks
message("The optimal K is ", Ks$K[which.min(Ks$AICc)])
```

<br>

**8.2.2 Model Performance**<br>
Table below summarizes the performance of the Dynamic Harmonic Regression model.
```{r}
# Model with optimal K
summary(fit_dhx1)

# tsCV
get_dhx1_rmse <- function(ts_train, ts_test) {
  dhx1_fit <- auto.arima(ts_train[,'Y'], xreg=cbind(ts_train[,c('Xc', 'Xd')], fourier(ts_train[,'Y'], K = 1)), seasonal=F)
  dhx1_fc <- forecast(dhx1_fit, xreg = cbind(ts_test[,c('Xc', 'Xd')], fourier(ts_train[,'Y'], K = 1, h=12)))
  dhx1_fci <- Y_inv(dhx1_fc)
  v_actual <- as.numeric(ts_test[,'sales_qty'])
  return(rmse(as.numeric(dhx1_fci$mean), v_actual))
}

get_dhx1i_rmse <- function(i) {
  ts_train <- head(ts10_train, i)
  ts_test <- ts10_train[(i+1):(i+12),]
  return(get_dhx1_rmse(ts_train, ts_test))
}

dhx110_is <- seq(nrow(ts10_train)/2, nrow(ts10_train)-12)
dhx110_tscv <- data.frame(i = dhx110_is, rmse = sapply(dhx110_is, get_dhx1i_rmse))

# Performance
fci_dhx1_rmse <- rmse(as.numeric(fci_dhx1$mean), ts10s_actual)
fci_dhx1_mase <- mase(as.numeric(fci_dhx1$mean), ts10s_actual)
dhx1_tscv_rmse <- round(mean(dhx110_tscv$rmse), digits=2)
message("Dynamic Harmonic model: MASE=", fci_dhx1_mase, " RMSE=", fci_dhx1_rmse, " RMSE(tsCV)=", dhx1_tscv_rmse)
```

<br>

