---
title: "ARIMA"
author: "Yang Xi"
date: "Created on 17 Jan 2018; Modified on 24 Aug 2021"
output:
  html_document:
    toc: 1
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

<br>

# 6. ARIMA
### 6.1 Definition
**6.1.1 Non-Seasonal ARIMA Model**<br>
*Autoregressive (AR) model*: multiple regression with lagged observations as predictors 
$$ y=c+\phi_1 y_{t-1}+\phi_2 y_{t-2}+⋯+\phi_p y_{t-p}+e_t $$
*Moving Average (MA) model*: Multiple regression with lagged errors as predictors
$$y=c+e_t+\theta_1 e_{t-1}+\theta_2 e_{t-2}+⋯+\theta_q e_{t-q} $$
Combining the AR and MA models results in *ARMA model*:
$$ y=c+\phi_1 y_{t-1}+\phi_2 y_{t-2}+⋯+\phi_p y_{t-p}+\theta_1 e_{t-1}+\theta_2 e_{t-2}+⋯+\theta_q e_{t-q}+e_t $$
$c$ is also called the drift component.

<br>
Integrate the ARMA models with differencing results in the __*ARIMA(p, d, q)*__ models,
where $d=\text{number of differencing}$

<br>

**6.1.2 Seasonal ARIMA Model**<br>
The seasonal ARIMA model is denoted as __*ARIMA(p, d, q)(P, D, Q)m*__, where
$$ D=\text{Number of seasonal differences} $$
$$ P=\text{Number of seasonal AR lags: }y_{t-m},y_{t-2m},…,y_{t-Pm} $$
$$ Q=\text{Number of seasonal MA lags: }e_{t-m},e_{t-2m},…,e_{t-Qm} $$
$$ m=\text{Number of observations per period} $$

<br>

### 6.2 Parameter Selection
The `auto.arima()` function from `forecast` package in R implements the Hyndman-Khandakar algorithm to select the optimal ARIMA parameters of the given data:

1. Select number of differencing d via unit root tests.
2. Select p and q by minimizing AICc.
3. Estimate parameters using maximum likelihood estimation.
4. Use stepwise search to traverse the model space. Stepwise search is used to save time, while it may not return global optimal.

<br>

### 6.3 Example 1: Non-Seasonal ARIMA of Monthly Sales Data
The same data used in the basic algorithms and exponential weighted forecast sections will be demonstrated in this section, which is the monthly sales data of an item:
```{r, message=FALSE}
library(dplyr)
library(fpp2)

dfItemC <- read.csv('data/time_series_monthly_C.csv')

tsC <- select(dfItemC, -first_date)
row.names(tsC) <-dfItemC$first_date
tsC <- ts(tsC, start=c(2008, 1), frequency=12)
tsCs <- tsC[,'sales_qty']
autoplot(tsCs)
ggAcf(tsCs)

tsCs_train <- window(tsCs, end = c(2013, 0))
tsCs_test <- window(tsCs, start = c(2013, 1))
tsCs_actual <- as.numeric(tsCs_test)
```
<br>
Recall that the 2008 – 2012 data is used as training set, and the 2013 data as test set.

<br>

**6.3.1	ARIMA Parameters**<br>
By applying `auto.arima()` to the training data, the following model is fitted:
```{r}
# Forecast Accuracy
rmse_e <- function(e) {
  round(sqrt(mean(e^2, na.rm=T)), digits=2)
}


rmse <- function(pred, actual) {
  rmse_e(pred-actual)
}

mase <- function(pred, actual) {
  l <- length(pred)
  round(sum(abs(pred-actual))/sum(abs(diff(actual)))*(l-1)/l, digits=2)
}

# Fit an automatic ARIMA model
fit_aa1 <- auto.arima(tsCs_train)
summary(fit_aa1)

# ARIMA(1,1,1) for comparison
#fit_a1 <- arima(tsCs_train, order=c(1,1,1))
#summary(fit_a1)
```
<br>
The optimal model is selected as a non-seasonal ARIMA model with p=4, q=1 integrated by 1 level of differencing.

<br>

**6.3.2 Forecast**<br>
The following figures present the forecast and residual plots of the model:
```{r}
fc_aa1 <- forecast(fit_aa1, h=12)
autoplot(fc_aa1, ylim=c(0, 3500))  + autolayer(tsCs_test)
checkresiduals(fit_aa1)
```

<br>

**6.3.3 Performance**<br>
Table below summarizes the performance of the ARIMA model:
```{r}
fc_aa1_rmse <- rmse(as.numeric(fc_aa1$mean), tsCs_actual)
fc_aa1_mase <- mase(as.numeric(fc_aa1$mean), tsCs_actual)

tscv_aa1 <- tsCV(tsCs_train, function(x,h) forecast(auto.arima(x),h=h), h = 12)
tscv_aa1_rmse <- rmse_e(as.numeric(tscv_aa1))
message("auto.arima model1: MASE=", fc_aa1_mase, " RMSE=", fc_aa1_rmse, " RMSE(tsCV)=", tscv_aa1_rmse)
```

<br>

### 6.4 Example 2: Seasonal ARIMA of Daily Sales Data
Daily sales data of an item from 2014 Q3 to 2016 Q2 will be used as example. Recall its time series plot and ACF plot:
```{r, message=FALSE}
dfItemB <- read.csv('data/time_series_daily_B.csv')

# Convert to ts
tsB <- select(dfItemB, -date)
row.names(tsB) <- dfItemB$date
tsB_p <- 91 # 91 days in one quarter (decimal frequency may result fail in seasonal model)
tsB <- ts(tsB, start=c(1, 1), frequency=tsB_p)
tsBs <- tsB[,'sales_qty']

autoplot(tsBs)
ggAcf(tsBs)

tsBs_train <- window(tsBs, end = c(8, 3))
tsBs_test <- window(tsBs, start = c(8, 4))
tsBs_actual <- as.numeric(tsBs_test)
```
<br>
2014 Q3 to 2016 Q1 (7 quarters) data will be used as training set, and 2016 Q2 data as test set.

<br>

**6.4.1	ARIMA Parameters**<br>
By applying `auto.arima()` to the training data, the following model is fitted:
```{r}
fit_aa2 <- auto.arima(tsBs_train)
summary(fit_aa2)

# ARIMA(1,0,0)(0,0,1)[91] for comparison
#fit_a2 <- arima(tsCs_train, order=c(1,0,0), seasonal=list(order=c(0,0,1)))
#summary(fit_a2)
```
<br>
The optimal model is selected as a seasonal ARIMA model with p=1 and Q=1 with m=91. Note that the value of m is **NOT** selected by the `auto.arima()` function, but a user input observed from the ACF plot.

<br>

**6.4.2 Forecast**<br>
The following figures present the forecast and residual plots of the model:

```{r}
fc_aa2 <- forecast(fit_aa2, h=91)
autoplot(fc_aa2)
checkresiduals(fit_aa2)
```
<br>
Note that the residuals have high ACF value, which implies that there are patterns unexplained by this model.

<br>

**6.4.3 Performance**<br>
Performance of the seasonal ARIMA model is summarized as below:
```{r}
# Performance
fc_aa2_rmse <- rmse(as.numeric(fc_aa2$mean), tsBs_actual)
fc_aa2_mase <- mase(as.numeric(fc_aa2$mean), tsBs_actual)

## The following tsCV will take very long
## tscv_aa2 <- tsCV(tsBs_train, function(x,h) forecast(auto.arima(x),h=h), h = 91)
## tscv_aa2_rmse <- rmse_e(as.numeric(tscv_aa2))
tscv_aa2_rmse <- 8.46
message("auto.arima model2: MASE=", fc_aa2_mase, " RMSE=", fc_aa2_rmse, " RMSE(tsCV)=", tscv_aa2_rmse)
```
<br>
Seasonal Naïve model is also fitted as a benchmark:
```{r}
fc_snaive2 <- snaive(tsBs_train, h = 91)
checkresiduals(fc_snaive2)
autoplot(fc_snaive2)

fc_snaive2_rmse <- rmse(as.numeric(fc_snaive2$mean), tsBs_actual)
fc_snaive2_mase <- mase(as.numeric(fc_snaive2$mean), tsBs_actual)

tscv_snaive2 <- tsCV(tsBs_train, snaive, h = 91)
tscv_snaive2_rmse <- rmse_e(as.numeric(tscv_snaive2))
message("snaive model: MASE=", fc_snaive2_mase, " RMSE=", fc_snaive2_rmse, " RMSE(tsCV)=", tscv_snaive2_rmse)
```

<br>
