---
title: "Linear Regression (R)"
author: "Yang Xi"
date: "21 Nov, 2018"
output:
  html_document:
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

# <br>Model Fitting and Interpretation
### Independent Numeric and Categorical variables
Note that, `lm()` is equivalent to `glm()` wtih `family = gaussian` (default)
```{r}
# Simulate data
set.seed(1)
n <- 9000
x1 <- rep(c('A', 'B', 'C'), n/3)[sample(n, n)]
x2 <- runif(n, min=0, max=2)
x3 <- runif(n, min=0, max=1)
e <- runif(n, min=-0.5, max=0.5)

b1_map <- c('A'=-2, 'B'=-1, 'C'=1) # 0, 1, 3
b2 <- 3
y <- b1_map[x1] + b2*x2 + e

df <- data.frame(x1=x1, x2=x2, x3=x3, y=y)

# Fit model
lmFit <- lm(y~., df)
summary(lmFit)

```

**<br>Response to highly correlated numeric variables**
```{r}
# Simulate data
set.seed(1)
n <- 10000
x1 <- runif(n, min=1, max=2)
x2 <- x1
e <- runif(n, min=-0.5, max=0.5)

b1 <- 2
b2 <- 3
y <- b1*x1+b2*x2+e

df <- data.frame(x1=x1, x2=x2, y=y)

# Fit model
lmFit <- lm(y~., df)
summary(lmFit)
```

### <br>Interaction: numeric variable interact with boolean variable
Take note how to interpret the result
```{r}
# Simulate data
set.seed(1)
n <- 10000
x1 <- rep(c('N', 'Y'), n/2)[sample(n, n)]
x2 <- runif(n, min=0, max=1)
e <- runif(n, min=-0.1, max=0.1)

bn <- -2
by <- 1
y <- ifelse(x1=='N', bn*x2, by*x2) + e

df <- data.frame(x1=x1, x2=x2, y=y)

# Fit model
lmFit <- lm(y~x1*x2, df)
summary(lmFit)
```

### <br>Polynomial Numeric Variable
Take note `raw = TRUE`
```{r}
set.seed(1)
n <- 10000
x1 <- runif(n, min=2, max=5)
e <- runif(n, min=-0.5, max=0.5)

b2 <- 3
b1 <- 2
y <- b2*x1^2 + b1*x1 + e

df <- data.frame(x1=x1, y=y)

# Fit model
lmFit <- lm(y~poly(x1, 2, raw = TRUE), df) # raw = FALSE (default) is orthogonal polynomails
summary(lmFit)
```

# <br>Example: Classification
```{r, message=FALSE}
library(tidyverse)
library(caret)
library(MLmetrics)
library(ROCR)

dfTrain0 <- read_csv("../0. Data/(2016 UCI) Credit Default/data_train.csv")
dfTest0 <- read_csv("../0. Data/(2016 UCI) Credit Default/data_test.csv")

PrepTrainTest <- function(df) {
  df %>%
    mutate(Education = relevel(as.factor(Education), ref="high school"),
           SepRepayment = relevel(as.factor(SepRepayment), ref="paid"),
           Default = relevel(as.factor(ifelse(Default==0, "N", "Y")), ref="N") # for classProbs=T
           )
}
dfTrain <- PrepTrainTest(dfTrain0)

yWeights <- ifelse(dfTrain$Default=="N", 1, sum(dfTrain$Default=="N")/sum(dfTrain$Default=="Y"))

# Fit model with CV (10-fold by default)
## note: quasibinomial is used here because with weight the class became non-integer.
## it yields the same result as binomial, but supresses the warning message
f1 <- function(data, lev = NULL, model = "F1") {
  # take note the positive is lev[2] in this example
  c(F1 = F1_Score(data$obs, data$pred, lev[2]))
}

lmFit <- train(Default ~ ., data=dfTrain, method='glm', family = quasibinomial,
               weights=yWeights,
               metric = "F1",
               trControl=trainControl(method="cv",
                                      summaryFunction=f1))
summary(lmFit)
```

### <br>Train performance
```{r}
probTrain <- predict(lmFit, type="prob")$"Y"
predTrain <- predict(lmFit)

cmTrain <- confusionMatrix(predTrain, dfTrain$Default, positive="Y")
cmTrain

perfTrain <- data.frame(F1 = round(cmTrain$byClass['F1'], 3),
                        AUC = round(performance(prediction(probTrain, dfTrain$Default),
                                                measure="auc")@y.values[[1]], 3))
row.names(perfTrain) <- NULL
perfTrain
```

### <br>Cross-Validation performance
```{r}
lmFit
```

### <br>Test performance
```{r}
dfTest <- PrepTrainTest(dfTest0)

predTest <- predict(lmFit, dfTest)

cmTest <- confusionMatrix(predTest, dfTest$Default, positive="Y")
cmTest

paste('Test F1 score is', round(cmTest$byClass['F1'], 3))
```

# <br>Appendix: test with optimized cutoff for highest F1
### <br>Train performance
```{r}
lmFitOp <- train(Default ~ ., data=dfTrain, method='glm', family = quasibinomial,
                 weights=yWeights,
                 trControl=trainControl(method="none"))

# get optimal cutoff
prob <- predict(lmFitOp, type="prob")["Y"]
perf <- performance(prediction(prob, dfTrain$Default), measure="f")
plot(perf)
opCutoff <- perf@x.values[[1]][which.max(perf@y.values[[1]])]
opF1 <- max(perf@y.values[[1]], na.rm=T)
print(paste("Optimal train F1", round(opF1,3), "achieved at cutoff =", round(opCutoff, 3)))

predTrain <- as.factor(ifelse(prob < opCutoff, "N", "Y"))
confusionMatrix(predTrain, dfTrain$Default, positive="Y")
```

### <br>Cross-Validation performance
This will be done manually
```{r}
validate_fold <- function(valiIndex) {
  foldTrain <- dfTrain[-valiIndex,]
  foldVali <- dfTrain[valiIndex,]
  foldWeights <- ifelse(foldTrain$Default=="N", 1,
                        sum(foldTrain$Default=="N")/sum(foldTrain$Default=="Y"))
  
  foldFit <- train(Default ~ ., data=foldTrain, method='glm', family = quasibinomial,
                   weights=foldWeights,
                   trControl=trainControl(method="none"))
  
  # get optimal cutoff
  prob <- predict(foldFit, type="prob")["Y"]
  perf <- performance(prediction(prob, foldTrain$Default), measure="f")
  opCutoff <- perf@x.values[[1]][which.max(perf@y.values[[1]])]
  
  # validateion performance
  ## optimal F1
  prob <- predict(foldFit, foldVali, type="prob")["Y"]
  predOp <- as.factor(ifelse(prob < opCutoff, "N", "Y"))
  opF1 <- confusionMatrix(predOp, foldVali$Default, positive="Y")$byClass['F1']
  
  ## normal F1
  pred <- predict(foldFit, foldVali)
  F1 <- confusionMatrix(pred, foldVali$Default, positive="Y")$byClass['F1']
  
  foldOutput <- data.frame(opCutoff = opCutoff, opF1 = opF1, F1 = F1)
  row.names(foldOutput) <- NULL
  return(foldOutput)
}

cvValiIndex <- createFolds(dfTrain$Default, k=10)
cvResults <- map_dfr(cvValiIndex, validate_fold)

cvResults
print(paste("Average optimized F1 =", round(mean(cvResults$opF1),3)))
print(paste("Average normal F1 =", round(mean(cvResults$F1),3)))
```

### <br>Test performance
```{r}
probTest <- predict(lmFitOp, dfTest, type="prob")["Y"]
predTest <-  as.factor(ifelse(probTest < opCutoff, "N", "Y"))

cmTest <- confusionMatrix(predTest, dfTest$Default, positive="Y")
cmTest

paste('Test F1 score is', round(cmTest$byClass['F1'], 3))
```

